Supporting Information

### 2015-08-17 manuscript.Rmd
Moved LAI-BA to SuppInfo; fixed SuppInfo so it is standalone; changed Fig 2 to LAI instead of LAI anomaly; changed fig labels to LAI

@@

## Data analysis
We used generalized additive models (GAM) for all analyses of LAI and $\tau_d$ timeseries [@Wood2006], using the *mgcv* package in R version `r r_version` [@RCT2015]. A GAM is a semi-parametric fitting method that allows the estimation of smoothed trends in the data, including confidence intervals. In all fitted GAMs, we used a cubic regression spline. For the smoothed term in the model (days since start of experiment), we used up to `r .gambase` degrees of freedom, which resulted in biologically realistic smoothed LAI dynamics. Within the GAM, we used a penalized regression smoother [@Wood2006] to allow the final degree of smoothness to be estimated from the data. Random effects were included to account for the grouping of data within FACE rings, except when we analysed average LAI across all rings (Fig. 7). For Fig. 7, we estimated the derivative of the fitted GAM, including confidence intervals for the derivative, following Simpson [-@Simpson2014; see also @Curtis2014].

@@

\
We found no differences in LAI between the C~a~ treatments, as shown by consistent overlap of the 95% confidence intervals of the generalized additive model (thus, p > 0.05 at any point in time, Fig. 3). A brief exception was the period where one of the rings was affected by an insect outbreak (see above). A linear mixed-effects model with C~a~, time and pre-treatment basal area (Fig. 3) as fixed effects, and FACE ring as a random effect also showed no significant effect of C~a~ on LAI (p **fix this**). 
** rethink the following; CI on least-square means accounting for BA?**
But what difference between C~a~ treatments could we have reasonably detected? The average half-width of the confidence interval on the LAI anomaly (Fig. 4) was `r f(LanomCIamb)` m^2^m^-2^ for aC~a~, and `r f(LanomCIele)` m^2^m^-2^ for the eC~a~ treatment. This means that, provided the within-treatment variance stays the same, we could detect a difference of `r f(LanomCIamb + LanomCIele)` m^2^ m^-2^ (the sum of the half-width CI's), or about a change of `r f(Ldetect)`% over the mean LAI of the EucFACE (`r f(mean(ba$LAI.mean))` m^2^m^-2^).

\
Litter production correlated significantly with $\Delta L_\tau$ when the latter was separated into periods where $\Delta L_\tau$ > 0 (net leaf growth) and $\Delta L_\tau$ <  0 (net leaf loss) (Fig. 4). During periods of net leaf loss, we found a significant correlation (with a slope of `r f(tidy(lmlitdown)$estimate[2])`, SE = `r f(tidy(lmlitdown)$std.error[2])`, n = `r glance(lmlitdown)$df.residual+2`), close to the 1:1 line, as expected since no leaf growth was observed during any period where LAI was decreasing. During periods of net leaf growth, we found a significant correlation (with a slope of `r f(tidy(lmlitup)$estimate[2])`, SE = `r f(tidy(lmlitup)$std.error[2])`, n = `r glance(lmlitup)$df.residual+2`) between $\Delta L_\tau$ and litter production (Fig. 5).

\
The temporal dynamics of litter and leaf production are shown in Fig. 5, and demonstrate the irregular temporal pattern in leaf production and P~L~. We found no significant difference in either litter or leaf production between C~a~ treatments at any time point (p > 0.2 for both, assessed with linear-mixed effects models).

\
Next, we related P~L~ to average $LAI_\tau$ (during the litter collection period) across the six FACE rings (Fig. 6a). If leaf lifespan is constant, P~L~ would be linearly related to $LAI_\tau$. Instead, we found that in eC~a~, P~L~ was lower at a given $LAI_\tau$ as compared to aC~a~ (Fig. 6a). We estimated mean leaf lifespan from Fig. 6a as the ratio of $LAI_\tau$ to annual total P~L~. We found a significant (p `r pval(LLtest$p.value)`) increase in leaf lifespan (LL) with elevated C~a~ (Fig. 6b), with LL = `r f(LLtest$estimate1)`yr in aC~a~ and `r f(LLtest$estimate2)`yr in eC~a~.

\
Finally, we quantified the relationship between timing of leaf flushing and environmental drivers. During the `r nummon` month study period, we observed four leaf flushing events in the field, with one before estimates of $LAI_\tau$ were available (Fig. 7). These events occurred in spring and summer, with no clear seasonal cycle (October 2012, January 2013, December 2013 and January 2015). The dynamics of $LAI_\tau$ from the sensor record matched these events observed in the field quite well (Fig. 7a), as estimated from a generalized additive model fitted to the average $LAI_\tau$ across the six FACE rings. From the fitted model, we estimated the numerical derivative of LAI (dL/dt), as well as its confidence interval. All four leaf flushing events corresponded to significant positive dL/dt. One brief significant positive dL/dt (June 2013) did not correspond to an observed leaf flushing event, but it was small in magnitude. It is also obvious that every leaf flush event occurred shortly after a significant rainfall event (Fig. 4 and 6), when the soil water content in the top 80 cm increased to ca. > 0.2 m^3^ m^-3^. Two periods of sustained decrease in LAI (red lines in Fig. 7b, during winter of 2013 and 2014) corresponded to periods of low soil water availability.

@@

## Response of LAI to elevated [CO~2~]
We report on `r nummon` months of continuous leaf area index (LAI) measurements in a native evergreen *Eucalyptus* woodland exposed to +150ppm elevated CO~2~ (eC~a~), including 3 months of ramp-up in C~a~ from November 2012 to January 2013. We predicted that LAI would increase in response to eC~a~ as a result of enhanced CO~2~ assimilation for the entire canopy and increased soil water availability. We also expected that this eucalypt woodland, with its inherently low LAI [@Whitehead2004], would be even more responsive to eC~a~ than higher-LAI and more closed-canopy stands [@Norby2011]. Counter to our expectations but similar to Norby et al. [-@Norby2003], we found no significant effect of elevated C~a~ on LAI. Because estimated leaf lifespan averaged `r round(mean(ba$LL),1)` yrs (Fig. 6), the canopy has regrown at least twice during the study period, giving ample opportunity for eC~a~ effects to occur. It could be argued that the change in LAI in response to eC~a~ was too small to be detected. Natural variation in LAI among plots in native woodland is indeed a challenge to manipulative studies with low replication, such as FACE experiments. To increase sensitivity and statistical robustness, we used a novel method to account for pre-treatment differences by using stand basal area as a covariate. Using this method, we estimated that we can detect a difference of only a `r format(Ldetect,digits=1)`% in LAI. Based on this small detectable difference, we conclude that the lack of a significant C~a~ effect is not due to a lack of power, but a true lack of response to elevated C~a~.

@@

\
A possible explanation for the lack of a LAI response to eC~a~ is that leaf growth may not be directly related to increased carbon availability as a result of eC~a~. In other words, stimulation of net leaf photosynthesis by eC~a~ (a common response, also observed at the EucFACE, Gimeno et al. *in review*) may not lead to increased net leaf growth, due to other constraints such as water or nutrient limitations [@Crous2015]. Net leaf growth is the net sum of allocation of photosynthates to leaf growth, and turnover of the leaf pool. Either process may itself be affected by eC~a~ [@DeKauwe2014]. We found a significant effect of eC~a~ on leaf lifespan, with leaves living approximately `r round(abs(LLtest$estimate*12),1)` months longer in elevated C~a~ (Fig. 6) (an increase of `r round(100*with(LLtest,estimate2/estimate1 - 1),0)`%). Further years of observation are needed to determine if such a difference is sustained. Because we found no difference in LAI with C~a~ treatment, this suggests a compensating difference in the fractional allocation of net primary production to leaf growth. To further understand the response of biomass allocation to eC~a~ at the EucFACE, all components of the carbon balance must be carefully quantified, a task underway but not yet complete.

\
Despite a lack of response of canopy LAI, we did find a significantly longer leaf lifespan (LL) in eC~a~. We could not determine whether there was an interaction between LL and water availability, since we estimated leaf turnover from mean LAI and total accumulated leaf litter (Fig. 6). Although LL was increased in eC~a~, this was not reflected in a difference in leaf litter production at any time point (Fig. 5), suggesting the change in LL reflects a small but significant shift in leaf retention, not related to decreased litter production during drought. There is mixed evidence from other FACE sites on the relationship between LL and eC~a~. Norby et al. [-@Norby2003] reported inconsistent effects of eC~a~ on leaf area duration in a deciduous *Liquidambar* plantation (ORNL-FACE). At the same site, Warren et al. [-@Warren2011] found increased senescence in the eC~a~ treatment in response to a combined drought and heat wave. McCarthy et al. [-@McCarthy2007] found no effects of eC~a~ on phenology in either *Pinus* or *Liquidambar* canopies at the DukeFACE, despite enhancement of LAI by eC~a~. Taylor et al. [-@Taylor2008] reported delayed autumn senescence in *Populus* in two FACE experiments (PopFACE and AspenFACE). Although this indicates the canopy stayed green longer in elevated C~a~, [@Gielen2001] reported faster turnover rate of leaves in elevated C~a~ at the PopFACE, leading to similar LAI in ambient C~a~ despite higher productivity. Li et al. [-@Li2000] reported delayed senescence of the oldest leaf cohort in evergreen *Quercus myrtifolia* exposed to eC~a~ in an open-top chamber experiment. Clearly, no consistent picture emerges from these comparisons.

@@

## Litter production and dynamics of LAI
We found a correlation between LAI dynamics and litter production, when the data were analyzed separately for periods of decreasing and increasing LAI (Fig. 4). As expected, during periods of decreasing LAI, the change in LAI was correlated with litter production, and we were able to use a four-month period in 2013 without significant rainfall to calibrate estimates of LAI against litter production. During periods of increasing LAI, litter fall increased sharply (Figs. 4 and 5), indicating replacement of old leaves by young leaves. Indeed we observed in the field that very few old leaves were present in the canopy after the significant flushing event of December 2013. This is in line with Pook [-@Pook1984], who showed in *Corymbia maculata* (a genus closely related to *Eucalyptus*) accelerated litter production during periods of net LAI increase, but with a lag between litter fall and production of about 2 months. Similarly, Stone and Bacon [-@Stone1995] showed that drought-stressed trees had higher litter production at the branch scale, but irrigated trees had larger litter production at the canopy scale. Since leaf flushing has been shown to be closely related to soil water availability in eucalypts [Fig. 7; @Pook1984a], net leaf turnover is driven by water availability both through increased litter production during extended droughts, and complete canopy replacement following leaf flushing (Fig. 5). The consequence is that leaf turnover is highly variable in time (Fig. 5).

## Consequences for phenology models
Current-generation phenology models are heavily biased towards temperate conditions and deciduous rather than evergreen species [@Primack2011], where leaf flushing occurs in spring in response to increase in temperature [@Richardson2012; @Savoy2015]. In contrast, phenology of evergreen *Eucalyptus* is more opportunistic, and shows rapid response to water availability (Fig. 7). Specht and Brouwer [-@Specht1975] demonstrated for four *Eucalyptus* species that leaf growth  peaked both in spring and autumn, with low winter temperatures and high summer temperatures (and low water availability) suppressing leaf growth. Similarly, Pook  [-@Pook1984a] demonstrated occurrence of leaf production during most months in *Corymbia maculata*, except during winter and suppression by dry periods. Thus a minimum temperature is also required for leaf growth to occur. We also have not observed leaf flushing in winter time, but at our site this was confounded with water availability as both winters during the study period were relatively dry (Fig. 7).  We argue that both minimum temperature and high soil moisture are required  for leaf growth of subtropical eucalypts such as *E. tereticornis*.

@@

```{r fig.width=4, fig.height=4, echo=FALSE, message=FALSE, eval=includeFigs}
figure2(ba)
```

**Figure 2**. Average $LAI_\tau$ over the entire study period plotted against the basal area (sum of stem cross-sectional areas) of the FACE rings ($R^2$ = `r glance(balm)$r.squared`, LAI = `r tidy(balm)$estimate[1]` + `r tidy(balm)$estimate[2]` BA). This relationship was used to calculate the LAI anomaly in Fig. 3, as the difference between LAI predicted from this regression and the actual $LAI_\tau$ at any given time. Red symbols are elevated C~a~, blue symbols ambient C~a~.

```{r fig.width=8, fig.height=4, echo=FALSE, message=FALSE, eval=includeFigs}
figure3(subset(facegap_cloudy_byring, Date < maxdate))
```

**Figure 3**.Time series of LAI after correcting for plot-to-plot differences in basal area (BA). At every time point, the LAI anomaly was calculated as $LAI_\tau$ minus LAI fitted from the relationship with BA shown in Fig. 2. Each symbol is a value for a FACE ring, but the smooth curve (fitted with a generalized additive model) takes into account the random plot effect. Note that the 95% confidence always overlap, denoting no significant treatment effect at $\alpha$ = 0.05, except for a brief period in late 2014. This difference was due to an insect attack that led to partial defoliation of the canopy in one FACE ring (see Fig. 1).

```{r fig.width=6, fig.height=4, echo=FALSE, message=FALSE, echo=FALSE, message=FALSE, eval=includeFigs}
figure4(dLAIlitter)
```

**Figure 4**. Comparison of absolute changes in LAI ($\Delta L$) to litter production (in units of LAI), calculated for each litter collection period (approx. monthly). To calculate $\Delta L$, we fit a generalized additive model fitted to data from each FACE ring separately, and estimated the change in L during the litter collection period (and expressed as monthly changes, by accounting for the differences in actual lengths of litter collection periods). Green symbols indicate periods during which $\Delta L$ was positive (net leaf production), orange symbols when $\Delta L$ was negative (net leaf loss). Solid black lines are 1:1 lines, coloured lines regression lines with 95% confidence intervals. The strong positive correlation between litter production and $\Delta L$ during periods of net leaf production indicates that during leaf flushing, old leaves are actively shed.

```{r fig.width=8, fig.height=4, echo=FALSE, message=FALSE, eval=includeFigs}
figure5(dLAIlitter)
```

**Figure 5**. Timeseries of litter production and estimated gross leaf production rate. Leaf production rate was estimated as the sum of the net change in LAI and litter production (converted to a per area basis using specific leaf area estimates, see Methods). Error bars are omitted for clarity. We did not find any differences in litter production or leaf production between C~a~ treatments (see Results).

```{r fig.width=8, fig.height=4, echo=FALSE, message=FALSE, eval=includeFigs}
figure6(ba)
```

**Figure 6**.  (a) Total litter production (converted from data in Fig. 5 to an annual estimate) for each of the FACE rings plotted against mean $LAI_\tau$ during the entire litter collection period. The ratio of mean $LAI_\tau$ to annual litter production gives the mean leaf life span. Grey lines indicate leaf lifespan isolines (in years). (b) Calculated leaf lifespan for the six FACE rings shows a significant difference between C~a~ treatments (t-test, p `r pval(LLtest$p.value)`).

```{r fig.width=10, fig.height=8, echo=FALSE, message=FALSE, eval=includeFigs}
figure7(subset(facegap_cloudy_byring, Date < maxdate),
               subset(facesoilwater, Date < maxdate),
               subset(faceraindaily, Date < maxdate),
               subset(airt, Date < maxdate))
```

**Figure 7**. Dynamics of LAI in relation to environmental drivers. (a) Smoothed time series of $LAI_\tau$ averaged across all six FACE rings. (b) From panel (a), we calculated the daily change in LAI (dL/dt), including the 95% confidence interval for the derivative of the time series. Green indicates significant positive dL/dt, red indicates significant negative dL/dt (at $\alpha$ = 0.05). Four of the five significant positive increases corresponded to approximate dates of commencement of leaf flushing (horizontal lines) observed in the field (except the third brief significant increase). (c) Time course of soil water content measured over 0-80 cm soil depth interval, averaged across all sensors and all rings (blue line), and daily rainfall (black lines).  (d) Average daily air temperature with a fitted smoother.

@@

>>>>

# Make figures
# source("make_figures.R")

@@

## Data analysis
We used generalized additive models (GAM) for all analyses of LAI and $\tau_d$ timeseries [@Wood2006], using the *mgcv* package in R version `r r_version` [@RCT2015]. A GAM is a semi-parametric fitting method that allows the estimation of smoothed trends in the data, including confidence intervals. In all fitted GAMs, we used a cubic regression spline. For the smoothed term in the model (days since start of experiment), we used up to `r .gambase` degrees of freedom, which resulted in biologically realistic smoothed LAI dynamics. Within the GAM, we used a penalized regression smoother [@Wood2006] to allow the final degree of smoothness to be estimated from the data. Random effects were included to account for the grouping of data within FACE rings, except when we analysed average LAI across all rings (Fig. 6). For Fig. 6, we estimated the derivative of the fitted GAM, including confidence intervals for the derivative, following Simpson [-@Simpson2014; see also @Curtis2014].

@@

\
We found no differences in LAI between the C~a~ treatments, as shown by consistent overlap of the 95% confidence intervals of the generalized additive model (thus, p > 0.05 at any point in time, Fig. 2). A linear mixed-effects model with C~a~, time and pre-treatment basal area (Fig. 3) as fixed effects, and FACE ring as a random effect also showed no significant effect of C~a~ on LAI (p **fix this**). 
** rethink the following; CI on least-square means accounting for BA?**
But what difference between C~a~ treatments could we have reasonably detected? The average half-width of the confidence interval on the LAI anomaly (Fig. 4) was `r f(LanomCIamb)` m^2^m^-2^ for aC~a~, and `r f(LanomCIele)` m^2^m^-2^ for the eC~a~ treatment. This means that, provided the within-treatment variance stays the same, we could detect a difference of `r f(LanomCIamb + LanomCIele)` m^2^ m^-2^ (the sum of the half-width CI's), or about a change of `r f(Ldetect)`% over the mean LAI of the EucFACE (`r f(mean(ba$LAI.mean))` m^2^m^-2^).

\
Litter production correlated significantly with $\Delta L_\tau$ when the latter was separated into periods where $\Delta L_\tau$ > 0 (net leaf growth) and $\Delta L_\tau$ <  0 (net leaf loss) (Fig. 3). During periods of net leaf loss, we found a significant correlation (with a slope of `r f(tidy(lmlitdown)$estimate[2])`, SE = `r f(tidy(lmlitdown)$std.error[2])`, n = `r glance(lmlitdown)$df.residual+2`), close to the 1:1 line, as expected since no leaf growth was observed during any period where LAI was decreasing. During periods of net leaf growth, we found a significant correlation (with a slope of `r f(tidy(lmlitup)$estimate[2])`, SE = `r f(tidy(lmlitup)$std.error[2])`, n = `r glance(lmlitup)$df.residual+2`) between $\Delta L_\tau$ and litter production (Fig. 4).

\
The temporal dynamics of litter and leaf production are shown in Fig. 4, and demonstrate the irregular temporal pattern in leaf production and P~L~. We found no significant difference in either litter or leaf production between C~a~ treatments at any time point (p > 0.2 for both, assessed with linear-mixed effects models).

\
Next, we related P~L~ to average $LAI_\tau$ (during the litter collection period) across the six FACE rings (Fig. 5a). If leaf lifespan is constant, P~L~ would be linearly related to $LAI_\tau$. Instead, we found that in eC~a~, P~L~ was lower at a given $LAI_\tau$ as compared to aC~a~ (Fig. 6a). We estimated mean leaf lifespan from Fig. 5a as the ratio of $LAI_\tau$ to annual total P~L~. We found a significant (p `r pval(LLtest$p.value)`) increase in leaf lifespan (LL) with elevated C~a~ (Fig. 5b), with LL = `r f(LLtest$estimate1)`yr in aC~a~ and `r f(LLtest$estimate2)`yr in eC~a~.

\
Finally, we quantified the relationship between timing of leaf flushing and environmental drivers. During the `r nummon` month study period, we observed four leaf flushing events in the field, with one before estimates of $LAI_\tau$ were available (Fig. 6). These events occurred in spring and summer, with no clear seasonal cycle (October 2012, January 2013, December 2013 and January 2015). The dynamics of $LAI_\tau$ from the sensor record matched these events observed in the field quite well (Fig. 6a), as estimated from a generalized additive model fitted to the average $LAI_\tau$ across the six FACE rings. From the fitted model, we estimated the numerical derivative of LAI (dL/dt), as well as its confidence interval. All four leaf flushing events corresponded to significant positive dL/dt. One brief significant positive dL/dt (June 2013) did not correspond to an observed leaf flushing event, but it was small in magnitude. It is also obvious that every leaf flush event occurred shortly after a significant rainfall event (Fig. 3 and 5), when the soil water content in the top 80 cm increased to ca. > 0.2 m^3^ m^-3^. Two periods of sustained decrease in LAI (red lines in Fig. 6b, during winter of 2013 and 2014) corresponded to periods of low soil water availability.

@@

## Response of LAI to elevated [CO~2~]
We report on `r nummon` months of continuous leaf area index (LAI) measurements in a native evergreen *Eucalyptus* woodland exposed to +150ppm elevated CO~2~ (eC~a~), including 3 months of ramp-up in C~a~ from November 2012 to January 2013. We predicted that LAI would increase in response to eC~a~ as a result of enhanced CO~2~ assimilation for the entire canopy and increased soil water availability. We also expected that this eucalypt woodland, with its inherently low LAI [@Whitehead2004], would be even more responsive to eC~a~ than higher-LAI and more closed-canopy stands [@Norby2011]. Counter to our expectations but similar to Norby et al. [-@Norby2003], we found no significant effect of elevated C~a~ on LAI. Because estimated leaf lifespan averaged `r round(mean(ba$LL),1)` yrs (Fig. 5), the canopy has regrown at least twice during the study period, giving ample opportunity for eC~a~ effects to occur. It could be argued that the change in LAI in response to eC~a~ was too small to be detected. Natural variation in LAI among plots in native woodland is indeed a challenge to manipulative studies with low replication, such as FACE experiments. To increase sensitivity and statistical robustness, we used a novel method to account for pre-treatment differences by using stand basal area as a covariate. Using this method, we estimated that we can detect a difference of only a `r format(Ldetect,digits=1)`% in LAI. Based on this small detectable difference, we conclude that the lack of a significant C~a~ effect is not due to a lack of power, but a true lack of response to elevated C~a~.

@@

\
A possible explanation for the lack of a LAI response to eC~a~ is that leaf growth may not be directly related to increased carbon availability as a result of eC~a~. In other words, stimulation of net leaf photosynthesis by eC~a~ (a common response, also observed at the EucFACE, Gimeno et al. *in review*) may not lead to increased net leaf growth, due to other constraints such as water or nutrient limitations [@Crous2015]. Net leaf growth is the net sum of allocation of photosynthates to leaf growth, and turnover of the leaf pool. Either process may itself be affected by eC~a~ [@DeKauwe2014]. We found a significant effect of eC~a~ on leaf lifespan, with leaves living approximately `r round(abs(LLtest$estimate*12),1)` months longer in elevated C~a~ (Fig. 5) (an increase of `r round(100*with(LLtest,estimate2/estimate1 - 1),0)`%). Further years of observation are needed to determine if such a difference is sustained. Because we found no difference in LAI with C~a~ treatment, this suggests a compensating difference in the fractional allocation of net primary production to leaf growth. To further understand the response of biomass allocation to eC~a~ at the EucFACE, all components of the carbon balance must be carefully quantified, a task underway but not yet complete.

\
Despite a lack of response of canopy LAI, we did find a significantly longer leaf lifespan (LL) in eC~a~. We could not determine whether there was an interaction between LL and water availability, since we estimated leaf turnover from mean LAI and total accumulated leaf litter (Fig. 5). Although LL was increased in eC~a~, this was not reflected in a difference in leaf litter production at any time point (Fig. 4), suggesting the change in LL reflects a small but significant shift in leaf retention, not related to decreased litter production during drought. There is mixed evidence from other FACE sites on the relationship between LL and eC~a~. Norby et al. [-@Norby2003] reported inconsistent effects of eC~a~ on leaf area duration in a deciduous *Liquidambar* plantation (ORNL-FACE). At the same site, Warren et al. [-@Warren2011] found increased senescence in the eC~a~ treatment in response to a combined drought and heat wave. McCarthy et al. [-@McCarthy2007] found no effects of eC~a~ on phenology in either *Pinus* or *Liquidambar* canopies at the DukeFACE, despite enhancement of LAI by eC~a~. Taylor et al. [-@Taylor2008] reported delayed autumn senescence in *Populus* in two FACE experiments (PopFACE and AspenFACE). Although this indicates the canopy stayed green longer in elevated C~a~, [@Gielen2001] reported faster turnover rate of leaves in elevated C~a~ at the PopFACE, leading to similar LAI in ambient C~a~ despite higher productivity. Li et al. [-@Li2000] reported delayed senescence of the oldest leaf cohort in evergreen *Quercus myrtifolia* exposed to eC~a~ in an open-top chamber experiment. Clearly, no consistent picture emerges from these comparisons.

@@

## Litter production and dynamics of LAI
We found a correlation between LAI dynamics and litter production, when the data were analyzed separately for periods of decreasing and increasing LAI (Fig. 3). As expected, during periods of decreasing LAI, the change in LAI was correlated with litter production, and we were able to use a four-month period in 2013 without significant rainfall to calibrate estimates of LAI against litter production. During periods of increasing LAI, litter fall increased sharply (Figs. 4 and 5), indicating replacement of old leaves by young leaves. Indeed we observed in the field that very few old leaves were present in the canopy after the significant flushing event of December 2013. This is in line with Pook [-@Pook1984], who showed in *Corymbia maculata* (a genus closely related to *Eucalyptus*) accelerated litter production during periods of net LAI increase, but with a lag between litter fall and production of about 2 months. Similarly, Stone and Bacon [-@Stone1995] showed that drought-stressed trees had higher litter production at the branch scale, but irrigated trees had larger litter production at the canopy scale. Since leaf flushing has been shown to be closely related to soil water availability in eucalypts [Fig. 6; @Pook1984a], net leaf turnover is driven by water availability both through increased litter production during extended droughts, and complete canopy replacement following leaf flushing (Fig. 4). The consequence is that leaf turnover is highly variable in time (Fig. 4).

## Consequences for phenology models
Current-generation phenology models are heavily biased towards temperate conditions and deciduous rather than evergreen species [@Primack2011], where leaf flushing occurs in spring in response to increase in temperature [@Richardson2012; @Savoy2015]. In contrast, phenology of evergreen *Eucalyptus* is more opportunistic, and shows rapid response to water availability (Fig. 6). Specht and Brouwer [-@Specht1975] demonstrated for four *Eucalyptus* species that leaf growth  peaked both in spring and autumn, with low winter temperatures and high summer temperatures (and low water availability) suppressing leaf growth. Similarly, Pook  [-@Pook1984a] demonstrated occurrence of leaf production during most months in *Corymbia maculata*, except during winter and suppression by dry periods. Thus a minimum temperature is also required for leaf growth to occur. We also have not observed leaf flushing in winter time, but at our site this was confounded with water availability as both winters during the study period were relatively dry (Fig. 6).  We argue that both minimum temperature and high soil moisture are required  for leaf growth of subtropical eucalypts such as *E. tereticornis*.

@@

```{r fig.width=8, fig.height=4, echo=FALSE, message=FALSE, eval=includeFigs}
figure2(subset(facegap_cloudy_byring, Date < maxdate))
```

**Figure 2**. Time series of LAI estimated from canopy transmittance. Each symbol is a value for a FACE ring, but the smooth curve (fitted with a generalized additive model) takes into account the random plot effect. Note that the 95% confidence always overlap, denoting no significant treatment effect at $\alpha$ = 0.05.

```{r fig.width=6, fig.height=4, echo=FALSE, message=FALSE, echo=FALSE, message=FALSE, eval=includeFigs}
figure3(dLAIlitter)
```

**Figure 3**. Comparison of absolute changes in LAI ($\Delta L$) to litter production (in units of LAI), calculated for each litter collection period (approx. monthly). To calculate $\Delta L$, we fit a generalized additive model fitted to data from each FACE ring separately, and estimated the change in L during the litter collection period (and expressed as monthly changes, by accounting for the differences in actual lengths of litter collection periods). Green symbols indicate periods during which $\Delta L$ was positive (net leaf production), orange symbols when $\Delta L$ was negative (net leaf loss). Solid black lines are 1:1 lines, coloured lines regression lines with 95% confidence intervals. The strong positive correlation between litter production and $\Delta L$ during periods of net leaf production indicates that during leaf flushing, old leaves are actively shed.

```{r fig.width=8, fig.height=4, echo=FALSE, message=FALSE, eval=includeFigs}
figure4(dLAIlitter)
```

**Figure 4**. Timeseries of litter production and estimated gross leaf production rate. Leaf production rate was estimated as the sum of the net change in LAI and litter production (converted to a per area basis using specific leaf area estimates, see Methods). Error bars are omitted for clarity. We did not find any differences in litter production or leaf production between C~a~ treatments (see Results).

```{r fig.width=8, fig.height=4, echo=FALSE, message=FALSE, eval=includeFigs}
figure5(ba)
```

**Figure 5**.  (a) Total litter production (converted from data in Fig. 5 to an annual estimate) for each of the FACE rings plotted against mean $LAI_\tau$ during the entire litter collection period. The ratio of mean $LAI_\tau$ to annual litter production gives the mean leaf life span. Grey lines indicate leaf lifespan isolines (in years). (b) Calculated leaf lifespan for the six FACE rings shows a significant difference between C~a~ treatments (t-test, p `r pval(LLtest$p.value)`).

```{r fig.width=10, fig.height=8, echo=FALSE, message=FALSE, eval=includeFigs}
figure6(subset(facegap_cloudy_byring, Date < maxdate),
               subset(facesoilwater, Date < maxdate),
               subset(faceraindaily, Date < maxdate),
               subset(airt, Date < maxdate))
```

**Figure 6**. Dynamics of LAI in relation to environmental drivers. (a) Smoothed time series of $LAI_\tau$ averaged across all six FACE rings. (b) From panel (a), we calculated the daily change in LAI (dL/dt), including the 95% confidence interval for the derivative of the time series. Green indicates significant positive dL/dt, red indicates significant negative dL/dt (at $\alpha$ = 0.05). Four of the five significant positive increases corresponded to approximate dates of commencement of leaf flushing (horizontal lines) observed in the field (except the third brief significant increase). (c) Time course of soil water content measured over 0-80 cm soil depth interval, averaged across all sensors and all rings (blue line), and daily rainfall (black lines).  (d) Average daily air temperature with a fitted smoother.

@@

**Figure S4**. Average $LAI_\tau$ over the entire study period plotted against the pre-treatment basal area (sum of stem cross-sectional areas) of the FACE rings.

### 2015-08-17 manuscript_supportinginfo.Rmd
Moved LAI-BA to SuppInfo; fixed SuppInfo so it is standalone; changed Fig 2 to LAI instead of LAI anomaly; changed fig labels to LAI

@@

@@

**Figure S2**. Correlation between LAI estimated from diffuse canopy transmittance ($L_\tau$) and from canopy photos (L~p~). Solid line is a 1:1 line, dashed line a regression line across all data (with the shading indicating the 95% confidence interval), and the colored lines regressions per plot across dates. (a) Estimates of L~p~ with the extinction coefficient (*k*) set to 0.5, as recommended by MacFarlane et al. [-@Macfarlane2007] ($L_\tau$ = `r tidy(lmcanphot)$estimate[1]` + `r tidy(lmcanphot)$estimate[2]` L~p~, R^2^ = `r glance(lmcanphot)$r.squared`, p `r pval(glance(lmcanphot)$p.value)`), (b) Estimates of L~p~ with *k* optimized to minimize the difference between L~p~ and $L_\tau$ ($L_\tau$ = `r tidy(lmcanphot2)$estimate[1]` + `r tidy(lmcanphot2)$estimate[2]`).

>>>>

```{r echo=FALSE, warning=FALSE, message=FALSE}
source("manuscript_calcs.R")
```

@@

**Figure S3**. Correlation between LAI estimated from diffuse canopy transmittance ($L_\tau$) and from canopy photos (L~p~). Solid line is a 1:1 line, dashed line a regression line across all data (with the shading indicating the 95% confidence interval), and the colored lines regressions per plot across dates. (a) Estimates of L~p~ with the extinction coefficient (*k*) set to 0.5, as recommended by MacFarlane et al. [-@Macfarlane2007] ($L_\tau$ = `r tidy(lmcanphot)$estimate[1]` + `r tidy(lmcanphot)$estimate[2]` L~p~, R^2^ = `r glance(lmcanphot)$r.squared`, p `r pval(glance(lmcanphot)$p.value)`), (b) Estimates of L~p~ with *k* optimized to minimize the difference between L~p~ and $L_\tau$ ($L_\tau$ = `r tidy(lmcanphot2)$estimate[1]` + `r tidy(lmcanphot2)$estimate[2]`).

\
```{r fig.width=4, fig.height=4, echo=FALSE, message=FALSE}
figureSI4(ba)
```

**Figure SI4**. Average $LAI_\tau$ over the entire study period plotted against the basal area (sum of stem cross-sectional areas) of the FACE rings ($R^2$ = `r glance(balm)$r.squared`, LAI = `r tidy(balm)$estimate[1]` + `r tidy(balm)$estimate[2]` BA). This relationship was used to calculate the LAI anomaly in Fig. 3, as the difference between LAI predicted from this regression and the actual $LAI_\tau$ at any given time. Red symbols are elevated C~a~, blue symbols ambient C~a~.

@@

## References
